Results

### 2021-08-04 Analysis/03_Summarise_results.Rmd
Update 03_Summarise_results.Rmd

```{r}
library(knitr)
## Loading the results
load("../Data/Processed/sim_equalizing.rda")
equalizing <- bkp
load("../Data/Processed/sim_facilitation.rda")
facilitation <- bkp
load("../Data/Processed/sim_filtering.rda")
filtering <- bkp
load("../Data/Processed/sim_competition.rda")
competition <- bkp
```

```{r}
## Loading the functions
source("../Functions/extract.table.R")
source("../Functions/plot.metric.results.R")
source("../Functions/paired.tests.R")
source("../Functions/split.metric.R")
source("../Functions/fit.model.rm.R")
source("../Functions/fancy.table.R")
source("../Functions/fancy.plot.R")
```

@@

```{r}
## Creating the list of results
all_results <- list(equalizing   = equalizing,
                    facilitation = facilitation,
                    filtering    = filtering,
                    competition  = competition)
## The different removal level names
removals <- c("20%", "40%", "60%", "80%")
## The colours
cols <- c("blue", "orange", "red", "green")
```

# TODOs:

  * [x]1. reworking the scaling
  * [x]2. plotting the metrics per stressors (i.e. all metrics for each stressor).
    `TODO: re-work the plots (15 rows * 4 columns; each row is a metric and contains the lines for the 4 rarefactions; each column is a stressor)`
  * [x]3. fitting a linear model per metric per stressor to rank models based on their r2
  * [x]4. t-test (paired) of difference between each random score and the stressor
  * [ ]5. run a psych-style pairwise comparison between each scaled metric
  * [ ]6. put the results in the context of the expectation table from the paper (Null vs process.docx)

@@

> TODO: we centre like this
> TODO: we scale like this

```{r}
## Extracting all the scaled results
scaled_results <- extract.table(all_results, scale.method = "between")
scaled_within <- extract.table(all_results, scale.method = "within")
```

@@

> TODO: we measure paired difference between random and stressor on the raw data
> TODO: we fitted a lm on the scaled data (scaled_metric ~ removal_level)

```{r}
## Slope and R2 for each model
model_fits <- lapply(scaled_results, fit.model.rm)
## Differences between stressor and random
null_differences <- lapply(all_results, paired.tests)

@@

>TG: Here is the visualisation of the results per stressor and per metric.

@@

>TG: Here is the visualisation of the results per stressor and per metric.

# My interpretation

 * The fact that most metrics pick up a negative change is interesting: a random decrease in number of elements often results in a smaller decrease in metric score than a non-random decrease in number of elements. In other words reducing X% of the elements in a biased way often reduces the score of all the metrics compared to a non-biased way. This is promising, if someone uses any metric to measure a decrease in something, it's better that it is due to a non-random cause rather than a random one ;).
 * The fact that different levels of removal (20%, 40%, etc...) results in different distributions of the metrics is also a good check (the level of the removal seems to often impact either the variance or the mean of the metric score - or both).
 * The fact that we can roughly see differences between the response of the metric and the different stressors is also encouraging (e.g. TPD divergence seems to pick up differences in scenarios pretty neatly!).
 * And finally, one that seems to be clearly working is the FD_evenness metric which seems to pick up changes in facilitation pretty neatly (if not too many species are removed).

>>>>

```{r, echo = FALSE}
library(knitr)
## Loading the results
load("../Data/Processed/sim_equalizing.rda")
equalizing <- bkp
load("../Data/Processed/sim_facilitation.rda")
facilitation <- bkp
load("../Data/Processed/sim_filtering.rda")
filtering <- bkp
load("../Data/Processed/sim_competition.rda")
competition <- bkp

## Loading the functions
source("../Functions/extract.table.R")
source("../Functions/plot.metric.results.R")
source("../Functions/paired.tests.R")
source("../Functions/split.metric.R")
source("../Functions/fit.model.rm.R")
source("../Functions/fancy.table.R")
source("../Functions/fancy.plot.R")
source("../Functions/fancy.names.R")
```

@@

```{r, echo = FALSE}
## Creating the list of results
all_results <- list(equalizing   = equalizing,
                    facilitation = facilitation,
                    filtering    = filtering,
                    competition  = competition)
## The different removal level names
removals <- c("20%", "40%", "60%", "80%")
## The colours
cols <- c("blue", "orange", "red", "green")
```

@@

We centre the metric based on the null results by substracting the score of the metric from the randomly reduced trait space to the stressed trait space (i.e. $score_{centred} = score_{stressed} - score_{random}$).
We then scale the centred metric by dividing it by it's maximum absolute centred value between all stressors (i.e. $score_{scaled} = score_{centred} / \text{max}(|scores_{centred}|))$).

```{r, echo = FALSE}
## Extracting all the scaled results
scaled_results <- extract.table(all_results, scale.method = "between")
scaled_within <- extract.table(all_results, scale.method = "within")
```

@@

To measure the differences between the random and stressed metric scores, we ran paired t-tests between each pairs of unscaled metric score for the random reduction of the trait space and the stressed reduction of the trait space (`t.test(random_scores, stressor_scores, paired = TRUE)`).

To check the trend in the different level of data removals (20, 40, 60 and 80%) using a linear model with the $scores_{scaled}$ as a response to the different removal levels (`lm(scaled_metric ~ removal_level`).

```{r, echo = FALSE}
## Slope and R2 for each model
model_fits <- lapply(scaled_results, fit.model.rm)
## Differences between stressor and random
null_differences <- lapply(all_results, paired.tests)

@@

@@

> I think that Figure X2 would do a good summary of how the metrics pick up changes. Here's a couple of interpretations of it in the light of my previous work on which metric to choose:
  * Don't use the TPD richness metric, it doesn't do much.
  * On the other hand, hypervolume regularity and richness are two good metrics because they only pick up really specific changes (respectively "equalizing" and "facilitation") and nothing else.

# Results interpretations in the light of our expectation table

These were our expectations:

Mechanism | Richness| Dispersion | Regularity
-------------|-------|-------|-------|
Equalizing   | Lower | Lower | Higher|
Facilitation | Higher| Higher| Higher|
Filtering (exclusion) | Higher| Higher| Higher|
Competition  | Lower | Lower |Nothing|

And these are our results per metric family:

 * [ok] prediction is correct (lower = decrease; higher = increase)
 * [-ok] prediction is kinda correct (there is a change compared to the null but not in the right direction)
 * [NO] to a wrong prediction: no change when change is expected or change when no change is expected

> NOTE: the difference between [ok] and [-ok] is actually not really meaningful here since these graphs shows differences compared to the null results so if both the null and stressor results increase but the stressor increases slowlier than the stressor, the results in the table will decrease. Although in general, if this is the case, this is a "bad" results since ideally we would want the null to not change at all (i.e. if the metric is just picking up changes in number of elements, this is not really useful).

#### Tree (dissimilarity)

Mechanism    | Richness (alpha)| Dispersion | Regularity
-------------|------------|-------------|-------|
Equalizing   | [ok] Lower | [ok] Lower  | [-ok] Higher|
Facilitation | [NO] Higher| [-ok] Higher| [-ok] Higher|
Exclusion    | [NO] Higher| [-ok] Higher| [-ok] Higher|
Competition  | [ok] Lower | [ok] Lower  | [NO] Nothing|

#### FD

Mechanism | Richness (Rao)| Dispersion | Regularity
-------------|-------|-------|-------|
Equalizing   | [NO] Lower  | [ok] Lower | [-ok] Higher|
Facilitation | [NO] Higher | [-ok] Higher| [NO] Higher|
Exclusion    | [-ok] Higher| [-ok] Higher| [-ok] Higher|
Competition  | [ok] Lower  | [ok] Lower | [NO] Nothing|

#### hypervolume

Mechanism | Richness (Rao)| Dispersion | Regularity
-------------|-------|-------|-------|
Equalizing   | [NO] Lower  | [-ok] Lower | [-ok] Higher|
Facilitation | [-ok] Higher| [NO] Higher| [NO] Higher|
Exclusion    | [NO] Higher | [NO] Higher| [NO] Higher|
Competition  | [ok] Lower  | [NO] Lower | [ok] Nothing|

#### TPD

Mechanism | Richness (Rao)| Dispersion | Regularity
-------------|-------|-------|-------|
Equalizing   | [NO] Lower  | [ok] Lower | [-ok] Higher|
Facilitation | [NO] Higher | [-ok] Higher| [-ok] Higher|
Exclusion    | [NO] Higher | [-ok] Higher| [-ok] Higher|
Competition  | [NO] Lower  | [ok] Lower | [NO] Nothing|

#### Other

I actually don't remember of the top of my head what the convex.hull and melodic rao/mpd are representing.

# Metrics correlations

And finally below are all the correlations per stressor between each metric per removal levels (20, 40, 60, 80%).

> TODO: do the PCA plots as suggested by Carlos to see things clearlier.

```{r, eval = TRUE, echo = FALSE, message = FALSE}
require(psych)

## Split the results per stressor and per metric
## (to extract the different removal levels)
splitted_results <- lapply(scaled_results, split.metric)

## Get the right removal level
get.matrix.level <- function(results, level = 1) {
    do.call(cbind, lapply(results, function(X, level) return(X[, level, drop = FALSE]), level = level))
}
## Wrapper for pairs.panels (with fancy names)
pairs.panels.wrapper <- function(results, ...) {
    ## Make fancy names
    colnames(results) <- fancy.names(colnames(results), sep = ":\n")
    psych::pairs.panels(results, ...)
}

```

## Equalizing

### Metrics correlation for equalizing with 20% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$equalizing, level = 1))
```

### Metrics correlation for equalizing with 40% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$equalizing, level = 2))
```

### Metrics correlation for equalizing with 60% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$equalizing, level = 3))
```

### Metrics correlation for equalizing with 80% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$equalizing, level = 4))
```

## Facilitation

### Metrics correlation for facilitation with 20% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$facilitation, level = 1))
```

### Metrics correlation for facilitation with 40% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$facilitation, level = 2))
```

### Metrics correlation for facilitation with 60% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$facilitation, level = 3))
```

### Metrics correlation for facilitation with 80% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$facilitation, level = 4))
```

## Filtering

### Metrics correlation for filtering with 20% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$filtering, level = 1))
```

### Metrics correlation for filtering with 40% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$filtering, level = 2))
```

### Metrics correlation for filtering with 60% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$filtering, level = 3))
```

### Metrics correlation for filtering with 80% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$filtering, level = 4))
```

## Competition

### Metrics correlation for competition with 20% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$competition, level = 1))
```

### Metrics correlation for competition with 40% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$competition, level = 2))
```

### Metrics correlation for competition with 60% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$competition, level = 3))
```

### Metrics correlation for competition with 80% removal

```{r, fig.height = 18, fig.width = 18, echo = FALSE, eval = TRUE}
## Equalizing level 1
pairs.panels.wrapper(get.matrix.level(splitted_results$competition, level = 4))
```
